Case based Arabic Morphological Analysis with Compositional Non-deterministic Automata

### 2010-06-25 tex/sarf.tex
spell check

\begin{abstract}
Natural language processing uses morphological 
analysis to abstract and annotate text.
Often times, and in particular in the context
of the Arabic language, annotations resulting from 
an ongoing branch of morphological analysis may not be
appropriate for the case study at hand. 
In this paper we present Sarf, a case based morphological 
analyser for Arabic text. 
Sarf allows a case-based controller to control and refine
the state of the analysis on the fly. 
It uses compositional non-deterministic finite state machines 
to analyse a stream of text. 
Each alive machine corresponds to a valid analysis. 
Sarf uses concatenative analysis based on recursive affixes 
to better preserve part of speech information.
We automated the analysis of three books of Islamic
narrations (hadith) using Sarf where we
abstracted each book into a vector of complex structures.
Our results show high accuracy and efficiency compared to
state of the art analyzers. 
\end{abstract}

@@

Automated analysis of Arabic data sets, including texts, 
publications, records and digital media is essential
with the huge digital Arabic content available nowadays. 
Morphological analysis is key in current automated 
analysis techniques for Arabic text. 
Current morphological analyzers~\cite{Sughaiyer:04}
use concatenative analysis
to analyze an Arabic word and
consider its internal structure composed of several
{\em morphemes}. A morpheme can be a {\em stem}, or an {\em affix}.
An affix can be a {\em prefix, suffix, } or an {\em infix}.
The analysis of one word may lead to several possible
morphological solutions.
\vocalize
For instance the word \RL{'a.hmadH} 
may have two valid morphological analyses where in the first
\RL{'a} is a prefix and the word means ``I praise him'',
and in the second \RL{'a} is part of the stem 
\RL{'a.hmad} (a proper noun)
and the word means ``his Ahmad''.

\novocalize
The solutions suffer from accuracy due to inherent difficulties
of morphological analysis of the Arabic language. 
For example, it is common practice to write Arabic text
without short vowels. 
Arabic letters can also have up to four different shapes 
corresponding to their position in a word, i.e, begining of 
a word, middle of a word, end of word, and separated. 
This allows two consecutive words such as 
\RL{almdrsT}\RL{al_A}
to be visually recognizable
as two separate words \RL{al_A} and \RL{almdrsT} without 
the need of a delimiter such as a space in between. 
This specifically happens when the first word ends with
a non-connecting letter and is referred to as the ``run-o'' words
problem~\cite{Buckwalter:04}.
Such forms occur in text and greatly increase the
difficulty of tokenization.

Current morphological analyzers such as 
Buckwalter~\shortcite{Buckwalter:02},
Xerox~\cite{Beesley:01}, SAMA~\cite{Kulick:10},
and ElixirFM~\cite{Otakar:07} exist.
They take as input white space delimited tokens~\cite{Kulick:10},
consider them as words,
and enumerate all possible solutions. 
This exhaustive enumeration may hurt performance and may
not be necessary or appropriate
in some case studies as noted in~\cite{Maamouri:10}. 
Moreover, a white space delimited token may have 
more than one word.
Other morphological analyzers such as Amira~\cite{Diab:07,Benajiba:07},
MAGEAD~\cite{Habash:05}, and MADA+TOKAN~\cite{Habash:09} 
also use machine learning and support vector machines (SVM) 
to enhance the accuracy of the morphological analysis at the expense 
of perfomance.

We hypothesise that many NLP case studies need the 
morphological analyser to answer simple queries that do not need 
high complexity and accuracy at the low morphological analysis level 
and that can often compensate for tolerable inaccuracy at higher level. 
For example, if the query is interested in proper names and the 
prefix in question in the analysis can only connect to a verb, 
we can provide an answer without completing the analysis. 
%TODO A lso if a word such as \RL{wasiim} was name was ambiguous and we were looking for a sequence of names, then 
We find evidence to our hypothesis in~\cite{Maamouri:10} where the 
addition of a new corpus to the Arabic Tree Bank~\cite{Maamouri:04}
with features demanding resolution at an abstraction level
higher than the text itself
led to a refined analyzer~\cite{Kulick:10}.  
We also find evidence in~\cite{Habash:06} that different types of 
morphological analyses behave differently on the same case study. 
We strongly beleive and we find evidence in our results that a 
case-based controller intervening at every decision is necessary to 
guide, use, prune, and refine the morphological analysis.

@@

In this paper, we present Sarf, a {\em novel case-based efficient
morphological analyzer} that uses parallel compositional 
non-deterministic automata driven by a case based controller.
Each alive machine in Sarf represents a valid morphological analysis. 
Sarf keeps alive all possible analyses of the text and gives 
opportunity to the controller to intervene at every single input 
character. 
The decisions of the case-based controller can thus prune false 
positives early in the run. 
Each alive machine in Sarf takes as input one character at a time 
from a text stream. 
Sarf does not assume that the word at hand is a token and
performs tokenization on the fly based on morphological correctness.
To out knowledge, this is the first morphological analyzer that 
handles the ``run-on'' words problem.

@@

We validate our hypothesis using a case study from the Islamic 
literature. 
A {\em hadith} is a narration related to the prophet Mohamad
through a {\em sanad} or a sequence of narrators. 
The authentication of a hadith higly depends on the credibility
of each of the narrators as reported in separate biography 
books. 
In this paper we consider the problem of automatically segmenting
a hadith book into narrations, then segmenting each narration into
its content or {\em matn} and its sanad.
We also would like to partition the sanad accurately into the 
separate narrators so that we can later look each one of them 
up in the biography books.

@@

We make the following key contributions. 
\begin{enumerate}
\item {\bf Case-based analysis:}  We deisgned and implemented
a case-based morphological analyzer where a case-based controller
machine controls and guides the analysis. 
\item {\bf Exhaustive analysis:} Since we have a case-based 
controller that can decide on the fly whether an analysis is 
accepted or not, we can afford to keep all valid analyses modulo
those pruned by the case controller. 
We do that using non-deterministic FSAs. 
\item {\bf Recursive affixes:}
We refine the concatenative analysis of 
Buckwalter~\shortcite{Buckwalter:02}  and use recursive affixes. 
This allowed Sarf to retain better compositional part of speech
tags.
\item {\bf Hadith case study:}
We evaluated Sarf using the hadith case study. We were able to
efficiently extract the sequence of narrators into a complex
data structure with three levels of hierarchy with high accuracy. 
\end{enumerate}

@@

Sarf is similar to Beesley~\shortcite{Beesley:01} in that it uses
finiste-state automata (FSA). 
Beesley reports that the number of the automata generated by a compiler
from Xerox rules is not controlled and also reports a difficulty to 
compose the FSAs into a single FSA framework. 
Sarf solves the problem with a non-deterministic composition of three
FSAs that are the prefix, stem, and suffix FSAs. 
With Sarf it is easy to control the FSAs with a user custom case 
controller, while with Beesley the effort requires recompiling the
Xerox rules.

Sarf is similar to The MADA+TOKAN toolkit in that it thinks of
the several valid morphological analyses as a richness that 
should be exploited at a higher abstract level rather than
an inaccuracy that should be corrected. 
Sarf differs in that it provides an interface for the 
case study to interfere in judging the analyses much earlier and
at a finer granularity level. 
Also in Sarf there is no need for a separate tokenizer such as
TOKAN as each alive automata carries with it the positions in 
text where a composition decision was made. 
The case-baded controller can build the trace when needed.

@@

We strongly feel that the refinemet of SAMA~\cite{Maamouri:10} 
is the closest to our approach. SAMA was refined to interact with
the ATB~\cite{Maamouri:04} project after the addition of a large 
new corpus. We find evidence in this decision that our hypothesis
about the utility of case-based analyzers. The refinements
that resulted in algorithmic changes in SAMA were done manually
and worked in integration with the ATB format only. 
In Sarf we propse a case-based controller approach that can modify 
the behaviour of the analyzer on the fly and that should interact
with the case study. In Sarf the analyzer remains general, and the 
case-based controller specializes the analyses.

@@

Like ElixirFM~\cite{Otakar:07} Sarf builds on the lexicon
of the Buckwalter analyzer. 
We also use deterministic parsing with tries or our analysis 
to run the prefix, suffix and stem FSAs. 
We think that their inferential-realizational approach 
that is highly compatible with the Arabic linguistic 
description~\cite{Badawi:04}
can make benefit of many features unique to the Arabic language.
Sarf leaves implementing that to the case-based controller
since in many cases the abstraction needed may not need the 
text to be grammar correct.
We can compare the case-based controller in Sarf to a high level
component in the ElixirFM domain sepcific language.

@@

Sarf encodes the prefix, suffix and stem lexicons into 
three linear FSAs and uses a non-deterministic composition
of the three FSAs to compute all valid morphological analyses.
We first present a runing example of Sarf. 
The running example explains
how the non-deterministic composition works and 
illustrates how Sarf implements recursive affixes and
handles ``run-on'' words.
Then we discuss how Sarf builds and composes the different FSAs.

@@

We start at the root square state in the prefix FSA
which is an accept state. 
The edge labeled $\epsilon$ that connects the prefix 
FSA to the stem FSA can transition from any accept state
in the prefix FSA to the root state in the stem FSA.
The edge that connects the stem FSA to the suffix FSA
follows the same behaviour.

When there are two valid transitions such as \RL{w} 
and $\epsilon$ as is the start case, a non-deterministic 
FSA $\Phi$ spawns an exact copy of itself $\Psi$. FSA $\Phi$
makes one transition, and FSA $\Psi$ makes the other. 
Each of the FSAs now represent a valid analysis so far. 
When the FSA is at a state where the next input character
leads to a reject state, the FSA dies.
The reject states are omitted from Figure~\ref{f:example}
and are denoted by the abscence of the corresponding 
characters. 
In our examples, if we assume there were no stems that start 
with the letter \RL{w}, $\Psi$ will die then. 
In reality probably $\Psi$ will die when the input string
is at \RL{wsyl`}.

@@

Note that $\Sigma$ can report a full word at any accept state
now by spawning a new automata using the $\epsilon$ transition
to the prefix FSA.
This is refined in Sarf to happen only with white space, delimiters, 
or non-connecting characters. 
We left that out in Figure~\ref{f:example} for clarity.

Finally, The case-based controller can interfere at any point in the 
analysis to make a decision like ignoring an automata that 
is not interesting. 
It may also decide to correct the analysis of one automata
based on the decision of the others.

@@

Many affixes are composed of other affixes. For example,
the prefix \RL{wsy} is composed of three other prefixes
namely \RL{w}, \RL{s}, and \RL{y}.
Moreover, the POS tag for \RL{wsy} is a concatenation
of the POS tags of each of the three morphemes. 
Keeping these morphemes separate in a list
is redundant and may produce consistency issues. 
We add two lists of compatibilty rules $L_{pp}$ and
$L_{xx}$ for prefix-prefix and suffix suffix compatibility
respectively.
This allows us to keep a shorter list of affixes. 
We need one more additional modification to accomodate 
a category for the morphemes accepted by the $L_{pp}$
and the $L_{xx}$ rules. For that we introduce the concept
of a {\em resulting category}. 
The resulting category in all practical cases
was the category of the second morpheme in the pair.

@@

We analyzed the prefixes and suffixes of 
Buckwalter~\cite{Buckwalter:02}
and extracted our refined lists of recursives affixes.
We encoded the lists into a FSA similar to that
in Figure~\ref{f:example}(a) and Figure~\ref{f:example}(c).
The two FSAs are directed acyclic graphs. 
The abscence of cycles in these FSAs reduces the 
traversal of the affix FSAs to a linear traversal.
This is computationaly superior to the 
approach of Buckwalter where the analyzer considers
all possible substrings of the word in question
and looks it up in the affix tables.

The prefix and suffix FSAs encode the $L_{pp}$ and
$L_{xx}$ compatibility lists.
An accept state in the FSA corresponds to the last character
of a prefix in the predix and suffix lists.
The accept states hold category state information as well
as POS and other tags.

@@

\subsection{Non-deterministic Composition of FSMs}
\label{sec:ndfsa}

@@

If $\Phi$ is in an accpet state and it is in the suffix
phase, then the algorithm checks whether it should report
a full word. 
This happens when $c$ is a white space or a delimiter, 
or when the last character leading to the current state
was a non-connecting character.

@@

\begin{table}[tb]
\centering
\begin{tabular} {p{6cm}}
\begin{Verbatim}[fontsize=\relsize{-1},
frame=topline,framesep=4mm,label=\fbox{NDSarf algorithm},
commandchars=\\\{\}, codes={\catcode`$=3\catcode`_=8}]
NDSarf(string $L$, Controller $A$) 
  FSAVec $M$; -- collection of FSA automata
  FSA $\Phi_1$;
  $M$.insert($\Phi_1$);
  foreach $c$ in $L$ \{
    foreach $\Phi$ in $M$ \{
      if ( $\Phi$.state.isAccept() ) \{
        if ($\Phi$.isSuffix())
          if ($c$.isWhite()) or 
            $A$.report();
          if ($\Phi$.lastChar().isNonConnecting())
            $A$.report();
        $\Psi$ = $\Phi$.clone();
        $M$.insert($\Psi$); \}

@@

A hadith in Islamic literature is a narration from the prophet Mohamad
related by multiple narrators.
Establishing the authenticity of a hadith is an important task
in Islamic studies. 
%Several researchers~\cite{Hadithopaedia:08} attempted to automate 
%the analysis of the hadith literature. 
In this paper we use Sarf to sucessfully automate the
analysis of three books of hadith selected 
arbitrarily~\cite{IbnHanbal,AlKulayni,AlTousi}\footnote{We obtained
  the digitized books from online sources such as 
  \href{http://www.yasoob.com/}{http://www.yasoob.com/} and 
  \href{http://www.al-eman.com/}{http://www.al-eman.com/}. 
  Our understanding is that those books are historical documents with 
  public IP.}.

Our analysis accpets a book as a text stream input
and segements it into a vector of individual hadiths. 
We segment each hadith into two parts. The first part
is the sanad and it
contains the chain of narrators who related the hadith
to Mohamad. 
The second part is the matn and it constitutes the content
of the hadith. 
We are concerned in further exploring the sanad. 
In particular, we detect the chain of narrators
and the relation between each narrator, his ancestor, and 
his predecessor in the chain. 
We also detect the full name of each narrator that is
usually composed of several proper names with connectors
in between.

>>>>

\begin{abstract}
Natural language processing uses morphological 
analysis to abstract and annotate text.
Often times, and in particular in the context
of the Arabic language, annotations resulting from 
an ongoing branch of morphological analysis may not be
appropriate for the case study at hand. 
In this paper we present Sarf, a case based morphological 
analyzer for Arabic text. 
Sarf allows a case-based controller to control and refine
the state of the analysis on the fly. 
It uses compositional non-deterministic finite state machines 
to analyze a stream of text. 
Each alive machine corresponds to a valid analysis. 
Sarf uses concatenative analysis based on recursive affixes 
to better preserve part of speech information.
We automated the analysis of three books of Islamic
narrations (hadith) using Sarf where we
abstracted each book into a vector of complex structures.
Our results show high accuracy and efficiency compared to
state of the art analyzers. 
\end{abstract}

@@

Automated analysis of Arabic data sets, including texts, 
publications, records and digital media is essential
with the huge digital Arabic content available nowadays. 
Morphological analysis is key in current automated 
analysis techniques for Arabic text. 
Current morphological analyzers~\cite{Sughaiyer:04}
use concatenative analysis
to analyze an Arabic word and
consider its internal structure composed of several
{\em morphemes}. A morpheme can be a {\em stem}, or an {\em affix}.
An affix can be a {\em prefix, suffix, } or an {\em infix}.
The analysis of one word may lead to several possible
morphological solutions.
\vocalize
For instance the word \RL{'a.hmadH} 
may have two valid morphological analyzes where in the first
\RL{'a} is a prefix and the word means ``I praise him'',
and in the second \RL{'a} is part of the stem 
\RL{'a.hmad} (a proper noun)
and the word means ``his Ahmad''.

\novocalize
The solutions suffer from accuracy due to inherent difficulties
of morphological analysis of the Arabic language. 
For example, it is common practice to write Arabic text
without short vowels. 
Arabic letters can also have up to four different shapes 
corresponding to their position in a word, i.e, beginning of 
a word, middle of a word, end of word, and separated. 
This allows two consecutive words such as 
\RL{almdrsT}\RL{al_A}
to be visually recognizable
as two separate words \RL{al_A} and \RL{almdrsT} without 
the need of a delimiter such as a space in between. 
This specifically happens when the first word ends with
a non-connecting letter and is referred to as the ``run-o'' words
problem~\cite{Buckwalter:04}.
Such forms occur in text and greatly increase the
difficulty of tokenization.

Current morphological analyzers such as 
Buckwalter~\shortcite{Buckwalter:02},
Xerox~\cite{Beesley:01}, SAMA~\cite{Kulick:10},
and ElixirFM~\cite{Otakar:07} exist.
They take as input white space delimited tokens~\cite{Kulick:10},
consider them as words,
and enumerate all possible solutions. 
This exhaustive enumeration may hurt performance and may
not be necessary or appropriate
in some case studies as noted in~\cite{Maamouri:10}. 
Moreover, a white space delimited token may have 
more than one word.
Other morphological analyzers such as Amira~\cite{Diab:07,Benajiba:07},
MAGEAD~\cite{Habash:05}, and MADA+TOKAN~\cite{Habash:09} 
also use machine learning and support vector machines (SVM) 
to enhance the accuracy of the morphological analysis at the expense 
of performance.

We hypothesize that many NLP case studies need the 
morphological analyzer to answer simple queries that do not need 
high complexity and accuracy at the low morphological analysis level 
and that can often compensate for tolerable inaccuracy at higher level. 
For example, if the query is interested in proper names and the 
prefix in question in the analysis can only connect to a verb, 
we can provide an answer without completing the analysis. 
%TODO A lso if a word such as \RL{wasiim} was name was ambiguous and we were looking for a sequence of names, then 
We find evidence to our hypothesis in~\cite{Maamouri:10} where the 
addition of a new corpus to the Arabic Tree Bank~\cite{Maamouri:04}
with features demanding resolution at an abstraction level
higher than the text itself
led to a refined analyzer~\cite{Kulick:10}.  
We also find evidence in~\cite{Habash:06} that different types of 
morphological analyses behave differently on the same case study. 
We strongly believe and we find evidence in our results that a 
case-based controller intervening at every decision is necessary to 
guide, use, prune, and refine the morphological analysis.

@@

In this paper, we present Sarf, a {\em novel case-based efficient
morphological analyzer} that uses parallel compositional 
non-deterministic Automata driven by a case based controller.
Each alive machine in Sarf represents a valid morphological analysis. 
Sarf keeps alive all possible analyses of the text and gives 
opportunity to the controller to intervene at every single input 
character. 
The decisions of the case-based controller can thus prune false 
positives early in the run. 
Each alive machine in Sarf takes as input one character at a time 
from a text stream. 
Sarf does not assume that the word at hand is a token and
performs tokenization on the fly based on morphological correctness.
To out knowledge, this is the first morphological analyzer that 
handles the ``run-on'' words problem.

@@

We validate our hypothesis using a case study from the Islamic 
literature. 
A {\em hadith} is a narration related to the prophet Mohammad
through a {\em sanad} or a sequence of narrators. 
The authentication of a hadith highly depends on the credibility
of each of the narrators as reported in separate biography 
books. 
In this paper we consider the problem of automatically segmenting
a hadith book into narrations, then segmenting each narration into
its content or {\em matn} and its sanad.
We also would like to partition the sanad accurately into the 
separate narrators so that we can later look each one of them 
up in the biography books.

@@

We make the following key contributions. 
\begin{enumerate}
\item {\bf Case-based analysis:}  We designed and implemented
a case-based morphological analyzer where a case-based controller
machine controls and guides the analysis. 
\item {\bf Exhaustive analysis:} Since we have a case-based 
controller that can decide on the fly whether an analysis is 
accepted or not, we can afford to keep all valid analyses modulo
those pruned by the case controller. 
We do that using non-deterministic FSAs. 
\item {\bf Recursive affixes:}
We refine the concatenative analysis of 
Buckwalter~\shortcite{Buckwalter:02}  and use recursive affixes. 
This allowed Sarf to retain better compositional part of speech
tags.
\item {\bf Hadith case study:}
We evaluated Sarf using the hadith case study. We were able to
efficiently extract the sequence of narrators into a complex
data structure with three levels of hierarchy with high accuracy. 
\end{enumerate}

@@

Sarf is similar to Beesley~\shortcite{Beesley:01} in that it uses
finite-state Automata (FSA). 
Beesley reports that the number of the Automata generated by a compiler
from Xerox rules is not controlled and also reports a difficulty to 
compose the FSAs into a single FSA framework. 
Sarf solves the problem with a non-deterministic composition of three
FSAs that are the prefix, stem, and suffix FSAs. 
With Sarf it is easy to control the FSAs with a user custom case 
controller, while with Beesley the effort requires recompiling the
Xerox rules.

Sarf is similar to The MADA+TOKAN toolkit in that it thinks of
the several valid morphological analyses as a richness that 
should be exploited at a higher abstract level rather than
an inaccuracy that should be corrected. 
Sarf differs in that it provides an interface for the 
case study to interfere in judging the analyses much earlier and
at a finer granularity level. 
Also in Sarf there is no need for a separate tokenizer such as
TOKAN as each alive Automata carries with it the positions in 
text where a composition decision was made. 
The case-based controller can build the trace when needed.

@@

We strongly feel that the refinement of SAMA~\cite{Maamouri:10} 
is the closest to our approach. SAMA was refined to interact with
the ATB~\cite{Maamouri:04} project after the addition of a large 
new corpus. We find evidence in this decision that our hypothesis
about the utility of case-based analyzers. The refinements
that resulted in algorithmic changes in SAMA were done manually
and worked in integration with the ATB format only. 
In Sarf we propose a case-based controller approach that can modify 
the behavior of the analyzer on the fly and that should interact
with the case study. In Sarf the analyzer remains general, and the 
case-based controller specializes the analyses.

@@

Like ElixirFM~\cite{Otakar:07} Sarf builds on the lexicon
of the Buckwalter analyzer. 
We also use deterministic parsing with tries or our analysis 
to run the prefix, suffix and stem FSAs. 
We think that their inferential-realizational approach 
that is highly compatible with the Arabic linguistic 
description~\cite{Badawi:04}
can make benefit of many features unique to the Arabic language.
Sarf leaves implementing that to the case-based controller
since in many cases the abstraction needed may not need the 
text to be grammar correct.
We can compare the case-based controller in Sarf to a high level
component in the ElixirFM domain specific language.

@@

Sarf encodes the prefix, suffix and stem lexicons into 
three linear FSAs and uses a non-deterministic composition
of the three FSAs to compute all valid morphological analyses.
We first present a running example of Sarf. 
The running example explains
how the non-deterministic composition works and 
illustrates how Sarf implements recursive affixes and
handles ``run-on'' words.
Then we discuss how Sarf builds and composes the different FSAs.

@@

We start at the root square state in the prefix FSA
which is an accept state. 
The edge labeled $\epsilon$ that connects the prefix 
FSA to the stem FSA can transition from any accept state
in the prefix FSA to the root state in the stem FSA.
The edge that connects the stem FSA to the suffix FSA
follows the same behavior.

When there are two valid transitions such as \RL{w} 
and $\epsilon$ as is the start case, a non-deterministic 
FSA $\Phi$ spawns an exact copy of itself $\Psi$. FSA $\Phi$
makes one transition, and FSA $\Psi$ makes the other. 
Each of the FSAs now represent a valid analysis so far. 
When the FSA is at a state where the next input character
leads to a reject state, the FSA dies.
The reject states are omitted from Figure~\ref{f:example}
and are denoted by the absence of the corresponding 
characters. 
In our examples, if we assume there were no stems that start 
with the letter \RL{w}, $\Psi$ will die then. 
In reality probably $\Psi$ will die when the input string
is at \RL{wsyl`}.

@@

Note that $\Sigma$ can report a full word at any accept state
now by spawning a new Automata using the $\epsilon$ transition
to the prefix FSA.
This is refined in Sarf to happen only with white space, delimiters, 
or non-connecting characters. 
We left that out in Figure~\ref{f:example} for clarity.

Finally, The case-based controller can interfere at any point in the 
analysis to make a decision like ignoring an Automata that 
is not interesting. 
It may also decide to correct the analysis of one Automata
based on the decision of the others.

@@

Many affixes are composed of other affixes. For example,
the prefix \RL{wsy} is composed of three other prefixes
namely \RL{w}, \RL{s}, and \RL{y}.
Moreover, the POS tag for \RL{wsy} is a concatenation
of the POS tags of each of the three morphemes. 
Keeping these morphemes separate in a list
is redundant and may produce consistency issues. 
We add two lists of compatibility rules $L_{pp}$ and
$L_{xx}$ for prefix-prefix and suffix suffix compatibility
respectively.
This allows us to keep a shorter list of affixes. 
We need one more additional modification to accommodate 
a category for the morphemes accepted by the $L_{pp}$
and the $L_{xx}$ rules. For that we introduce the concept
of a {\em resulting category}. 
The resulting category in all practical cases
was the category of the second morpheme in the pair.

@@

We analyzed the prefixes and suffixes of 
Buckwalter~\cite{Buckwalter:02}
and extracted our refined lists of recursive affixes.
We encoded the lists into a FSA similar to that
in Figure~\ref{f:example}(a) and Figure~\ref{f:example}(c).
The two FSAs are directed acyclic graphs. 
The absence of cycles in these FSAs reduces the 
traversal of the affix FSAs to a linear traversal.
This is computationally superior to the 
approach of Buckwalter where the analyzer considers
all possible substrings of the word in question
and looks it up in the affix tables.

The prefix and suffix FSAs encode the $L_{pp}$ and
$L_{xx}$ compatibility lists.
An accept state in the FSA corresponds to the last character
of a prefix in the prefix and suffix lists.
The accept states hold category state information as well
as POS and other tags.

@@

\subsection{Non-deterministic Composition of FSAs}
\label{sec:ndfsa}

@@

If $\Phi$ is in an accept state and it is in the suffix
phase, then the algorithm checks whether it should report
a full word. 
This happens when $c$ is a white space or a delimiter, 
or when the last character leading to the current state
was a non-connecting character.

@@

\begin{table}[tb]
\centering
\begin{tabular} {p{6cm}}
\begin{Verbatim}[fontsize=\relsize{-1},
frame=topline,framesep=4mm,label=\fbox{NDSarf algorithm},
commandchars=\\\{\}, codes={\catcode`$=3\catcode`_=8}]
NDSarf(string $L$, Controller $A$) 
  FSAVec $M$; -- collection of FSA Automata
  FSA $\Phi_1$;
  $M$.insert($\Phi_1$);
  foreach $c$ in $L$ \{
    foreach $\Phi$ in $M$ \{
      if ( $\Phi$.state.isAccept() ) \{
        if ($\Phi$.isSuffix())
          if ($c$.isWhite()) or 
            $A$.report();
          if ($\Phi$.lastChar().isNonConnecting())
            $A$.report();
        $\Psi$ = $\Phi$.clone();
        $M$.insert($\Psi$); \}

@@

A hadith in Islamic literature is a narration from the prophet Mohammad
related by multiple narrators.
Establishing the authenticity of a hadith is an important task
in Islamic studies. 
%Several researchers~\cite{Hadithopaedia:08} attempted to automate 
%the analysis of the hadith literature. 
In this paper we use Sarf to successfully automate the
analysis of three books of hadith selected 
arbitrarily~\cite{IbnHanbal,AlKulayni,AlTousi}\footnote{We obtained
  the digitized books from online sources such as 
  \href{http://www.yasoob.com/}{http://www.yasoob.com/} and 
  \href{http://www.al-eman.com/}{http://www.al-eman.com/}. 
  Our understanding is that those books are historical documents with 
  public IP.}.

Our analysis accepts a book as a text stream input
and segments it into a vector of individual hadiths. 
We segment each hadith into two parts. The first part
is the sanad and it
contains the chain of narrators who related the hadith
to Mohammad. 
The second part is the matn and it constitutes the content
of the hadith. 
We are concerned in further exploring the sanad. 
In particular, we detect the chain of narrators
and the relation between each narrator, his ancestor, and 
his predecessor in the chain. 
We also detect the full name of each narrator that is
usually composed of several proper names with connectors
in between.
